### 2021-05-04 written/eLifeRevision/Response.tex

\textbf{Reviewers were broadly enthusiastic about the potential usefulness of this methodology. However, the reviews and ensuing discussion revealed several points of concern about the manuscript and the approach. The full reviewer comments are included below.}

We thank the reviewers for the excellent and constructive feedback.  These valid criticisms of the original manuscript impressed upon us the importance of making key improvements to this research project.  We spent more than half of a year advancing this work according to the reviewer requests and suggestions, and we are excited to present them to you in a dramatically improved revision.

We have made significant changes to our writing and explanation of emergent property inference (EPI), done extensive comparisons of EPI to other parameter inference methods, and have seriously increased the quality and depth of our scientific analyses, which now yield strong theoretical results.  Our model analyses now focus on analyzing the rich structure of parameter distributions, which the deep probability distributions of EPI make possible.  Here, we provide a list of key manuscript updates, and below we explain how these changes address each of the reviewer’s concerns.

List of key manuscript improvements:
\begin{itemize}
\item We completely overhauled our presentation of the parameter inference technique emergent property inference (EPI). See response to main concern \#1.
\item We added an entirely new results section to compare EPI to alternative parameter inference techniques (SMC-ABC and SNPE) as parameter count increases, which demonstrates the advantages of EPI in high dimensional parameter spaces. See response to main concern \#2.
\item EPI is used to understand how noise across neuron type populations governs excitatory variability in a model of primary visual cortex.  
Here, EPI yields visually striking, novel insight where the conventional analytic approach became infeasible with increasing neuron types.  See response to main concern \#4.
\item EPI is used to discover multiple parametric regimes of rapid task switching in a model of superior colliculus.  
These regimes are efficiently identified and mechanistically characterized using the probabilistic modeling tools afforded by EPI, and we relate the inferred connectivities to results from optogenetic silencing experiments in rats. See response to main concern \#4.
\end{itemize}

@@

\begin{displayquote}
Lines 28-36 \\
``\quoting{intro_first}"
\end{displayquote}

\textbf{Other concerns: }

\textbf{(1) Maximizing the entropy of the distribution is not a reparameterization invariant exercise. That is, results depend on whether the model parameters contains rates or time constants, for example. I wonder if this approach is attempting to use a 'flat prior' in some sense which has the same reparameterization issue? Can the authors comment? }

The reviewer is correct to point out that maximum entropy solutions are not reparameterization invariant, and indeed the units matter.  The reviewer’s suggestion that the method is in some sense using a flat prior is also correct.  To clarify, EPI does not execute posterior inference, because there is no explicit dataset or specified prior belief in the EPI framework.  However, we derive the relation of EPI to bayesian inference in Section 5.1.6, which shows EPI uses a uniform prior when framed as variational inference.

@@

\begin{displayquote}
Lines 986-989 \\
``\quoting{ME_prior}"
\end{displayquote}

In our examples, we only infer distributions of parameters with the same units, so this issue should not draw concern over the validity of our model analyses.  As suggested, the EPI solution will differ according to relative scaling of parameter values under the maximum entropy selection principle. Thus, an important clarification is that sensitivity quantifications are made in the context of the chosen parameter scalings. 
%A final concern here is numerical: if the inferred distribution is in a thin region of parameter space, that is not well represented by the precision of the numerical format, there can be issues in optimization.  
%It makes most sense to make sure EPI is exploring through parameter space through an industry-recommended range of values, and that sensitivity measurements are interpretable.

@@

\textbf{This is a very interesting approach to an extremely important question in theoretical neuroscience, and the mathematics and algorithms appear to be very rigorous. The complexities in using this in practice make me wonder if it will find wide application though: setting up the objective to be differentiable, tweaking hyperparameters for training, and interpreting the results; all seem to require a lot from the user. On the other hand, the authors are to be congratulated on providing high quality open source code including clear tutorials on how to use it. }

\textbf{Major points }

\textbf{1. Training deep networks is hard. Indeed the authors devote a substantial amount of the manuscript to techniques for training them, and note that different hyperparameters were necessary for each of the different studies. Can the authors be confident that they have found the network which gives maximum entropy or close to it? If so, how. If not, how does that affect the conclusions? }

We thank the reviewer for their positive comments.  
To draw a parallel, training deep networks for visual processing used to be considered infeasible, but became easier through iterative improvements in architectural and hyperparameter choices that spread across the field.  
Similarly, training deep networks via EPI to learn parameter distributions became easier throughout this research project as we learned through trial and error what works well.  
In fact, there has been extraordinary progress in the field of deep probability distributions (specifically normalizing flows), that have allowed EPI to converge regularly while capturing complex structure (e.g. Dinh et al. 2017 and Kingma et al. 2018).  
This manuscript's explanation of hyperparameter choices and the extensive set of examples in the online code will serve as valuable guidelines for future research using this method.
Every figure of this paper is reproducible with the jupyter notebooks, and there are several tutorials for understanding the most consequential hyperparameters: the augmented Lagrangian constant and deep probability distribution architecture.

In general, we cannot know if we have obtained the global maximum entropy distribution for a given emergent property.  
The reviewer is correct to point out that multiple distributions may satisfy the emergent property and have different levels of entropy.  
In the new manuscript, we present the method as an inference technique without focusing very greatly on maximum entropy, since it tends to distract and confuse the reader.  
We derive an analytic equivalence to variational inference (Section 5.1.6) showing that a.) EPI is a valid inference method, and b.) to emphasize that maximum entropy is the normative selection principle of bayesian inference methods in general. 
Thus, the concern of not having the globally optimal inferred distribution is the same that applies to all other statistical inference techniques.

### 2021-05-04 written/eLifeRevision/Bittner_eLife2020.tex
through 3.4

\linenumbers
\section{Abstract}
A cornerstone of theoretical neuroscience is the circuit model: a system of equations that captures a hypothesized neural mechanism.  
Such models are valuable when they give rise to an experimentally observed phenomenon -- whether behavioral or a pattern of neural activity -- and thus can offer insights into neural computation.
The operation of these circuits, like all models, critically depends on the choices of model parameters.
A key step is then to identify the model parameters consistent with observed phenomena: to solve the inverse problem.
To solve challenging inverse problems concerning neural datasets, neuroscientists have used statistical inference techniques to much success. 
However, when theorizing circuit models, scientists predominantly focus on reproducing computational properties rather than explaining a given neural dataset well.
In this work, we present a novel technique, emergent property inference (EPI), that brings the power and versatility of the modern probabilistic modeling toolkit to theoretical neuroscience.
Our method uses deep neural networks to learn parameter distributions with complex structure that produce specific computational properties in circuit models.
This methodology is introduced through a motivational example inferring conductance parameters in a circuit model of the stomatogastric ganglion.
Then, with recurrent neural networks of increasing size, we show that EPI allows precise control over the behavior of inferred parameters, and that EPI scales better in parameter dimension than alternative techniques.
In the remainder of this work, we present novel theoretical findings gained through the examination of complex parametric structure captured by EPI.
In a model of primary visual cortex, we discovered how connectivity with multiple inhibitory subtypes shapes variability in the excitatory population.
Finally, in a model of superior colliculus, we identified and characterized two distinct regimes of connectivity that facilitate switching between opposite tasks amidst interleaved trials, mechanistically characterized each regime using probabilistic tools afforded by EPI, and found conditions where these circuit models reproduce results from optogenetic silencing experiments.
Beyond its scientific contribution, this work illustrates the variety of analyses possible once deep learning is harnessed towards solving theoretical inverse problems.

\section{Introduction}
\q{intro_first}{The fundamental practice of theoretical neuroscience is to use a mathematical model to understand neural computation, whether that computation enables perception, action, or some intermediate processing.  
A neural circuit is systematized with a set of equations -- the model -- and these equations are motivated by biophysics, neurophysiology, and other conceptual considerations \cite{kopell1988coupled,  marder1998biophysics, abbott2008theoretical, wang2010neurophysiological, o2015computational}.
The function of this system is governed by the choice of model \emph{parameters}, which when configured in a particular way, give rise to a measurable signature of a computation.   
The work of analyzing a model then requires solving the inverse problem: given a computation of interest, how can we reason about particular parameter configurations?  
The inverse problem is crucial for reasoning about likely parameter values, uniquenesses and degeneracies, and predictions made by the model \cite{gutenkunst2007universally, erguler2011practical, mannakee2016sloppiness}.}

@@

\q{intro_data_ep}{Statistical inference, of course, requires quantification of the vague term \emph{computation}.  
In neuroscience, two perspectives are dominant.  
First, often we directly use an \emph{exemplar dataset}: a collection of samples that express the computation of interest, this data being gathered either experimentally in the lab or from a computer simulation.  
While in some sense the best choice given its connection to experiment \cite{paninski2018neural}, some drawbacks exist: these data are well known to have features irrelevant to the computation of interest \cite{niell2010modulation, saleem2013integration, musall2019single}, confounding inferences made on such data.  
Related to this point, use of a conventional dataset encourages conventional data likelihoods or loss functions, which focus on some global metric like squared error or evidence, rather than the computation itself.  
Alternatively, researchers often quantify an \emph{emergent property}: a statistic of data that directly quantifies the computation of interest, wherein the dataset is only implicit.  
While such a choice may seem esoteric, it is not: the above ``gold standard" examples all quantify and focus on some derived feature of the data, rather than the data drawn from the model.}

\q{intro_epi}{An emergent property (EP) is of course a dataset by another name, but it suggests different approach to solving the same inverse problem:  here we directly specify the desired emergent property -- a statistic of datasets drawn from the model -- and the value we wish that property to have, and we set up an optimization program to find the distribution of parameters that produce this computation.  
This statistical framework is not new: it is intimately connected to the literature on approximate bayesian computation \cite{beaumont2002approximate, marjoram2003markov,sisson2007sequential}, parameter sensitivity analyses \cite{raue2009structural, karlsson2012efficient, hines2014determination, raman2017delineating}, maximum entropy modeling \cite{elsayed2017structure, savin2017maximum, mlynarski2020statistical}, and approximate bayesian inference \cite{tran2017hierarchical, gonccalves2019training}; we detail these connections in Section \ref{methods_related}.  
However, the adaptation of modern techniques to the problem of theoretical circuit analysis requires recent developments in deep learning for  constrained optimization \cite{loaiza2017maximum}, and architectural  choices for scalable, flexible generative modeling \cite{dinh2017density, kingma2018glow}.
We detail our method, which we call emergent property inference (EPI) in Section \ref{results_dgm}.}

Equipped with this method, we prove out EPI by demonstrating its capabilities and presenting novel theoretical findings from its analysis.
\q{intro_results1}{First, we show EPI's ability to handle biologically realistic circuit models using a five-neuron model of the stomatogastric ganglion \cite{gutierrez2013multiple}: a neural circuit whose parametric degeneracy is closely studied \cite{goldman2001global}.
Then, we show EPI's scalability to high dimensional parameter distributions by inferring connectivities of recurrent neural networks (RNNs) that exhibit stable, yet amplified responses -- a hallmark of neural responses throughout the brain \cite{murphy2009balanced, hennequin2014optimal, bondanelli2019population}.}
\q{intro_results2}{In a model of primary visual cortex \cite{litwin2016inhibitory, palmigiano2020structure}, EPI reveals the how recurrent processing across different neuron-type populations shapes excitatory variability: a finding that we show is analytically intractable.
Finally, we investigated the possible connectivities of superior colliculus that allow execution of different tasks on interleaved trials \cite{duan2018collicular}.
EPI discovered a rich distribution containing two connectivity regimes with different solution classes.
We queried the deep probability distribution learned by EPI to produce a mechanistic understanding of cortical responses in each regime. 
Intriguingly, the inferred connectivities of each regime reproduced results from optogenetic inactivation experiments in markedly different ways.}
These theoretical insights afforded by EPI illustrate the value of deep inference for the interrogation of neural circuit models.

@@

Second, once the model is selected, one must specify what the model should produce.
In this STG model, we are concerned with neural spiking frequency, which emerges from the dynamics of the circuit model (Fig. \ref{fig:STG}B).
An emergent property studied by Guttierez et al. of this stochastic model is the hub neuron firing at an intermediate frequency between the intrinsic spiking rates of the fast and slow populations.
This emergent property is shown in Figure \ref{fig:STG}C at an average frequency of 0.55Hz.
Our notion of intermediate hub frequency is not strictly 0.55Hz, but also moderate deviations of this frequency between the fast (.35Hz) and slow (.68Hz) frequencies, which are quantified in the emergent property with variance {0.025$^2$Hz$^2$}.
%To continue our running STG example, one such emergent property is the phenomenon of \emph{network syncing} -- in certain parameter regimes, the frequency of the hub neuron matches that of the fast and slow populations at an intermediate frequency.

%Third, qualitative parameter analysis ensues: since precise mathematical analysis is intractable in this model, a brute force sweep of parameters is done \cite{gutierrez2013multiple}.
%Subsequently, a qualitative description is formulated to describe the different parameter configurations that lead to the emergent property.  
%In this last step lies the opportunity for a precise quantification of the emergent property as a statistical feature of the model. 
% Once we have such a methodology, we can infer a probability distribution over parameter configurations that produce this emergent property.
%Sensitivity analyses \cite{raue2009structural} are also used to identify which parameter contributions have the greatest effect.
%In this last step lies the opportunity for a precise quantification of the emergent property as a statistical feature of the model.
Third, the model parameters producing the emergent property are inferred.
To infer the STG parameters of intermediate hub frequency with existing methodology, we need an exemplar dataset: experimentally recorded or synthesized.
By precisely quantifying the emergent property of interest as a statistical feature of the model, we use EPI to condition directly on this emergent property.
EPI learns a probability distribution of model parameters constrained to produce the emergent property.
In this last step lies the opportunity for a shift away from a dataset-oriented representation of model output towards that of an implicit dataset, where the only structure is the emergent property of interest.

\begin{figure}
\begin{center}
\includegraphics[scale=0.8]{figures/fig1/fig1.pdf}
\end{center}
\caption{Emergent property inference (EPI) in the stomatogastric ganglion.  
\textbf{A}. Conductance-based subcircuit model of the STG.
\textbf{B}. Spiking frequency $\omega(\mathbf{x}; \mathbf{z})$ is an emergent property statistic. 
Simulated at $g_{\text{el}} = 4.5$nS and $g_{\text{synA}} = 3$nS.
\textbf{C}. The emergent property of intermediate hub frequency.
Simulated activity traces are colored by $\log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ of generating parameters. (Panel E).
\textbf{D}. For a choice of circuit model and emergent property, emergent property inference (EPI) learns a deep probability distribution of parameters $\mathbf{z}$.
\textbf{E}. The EPI distribution producing intermediate hub frequency.
Samples are colored by log probability density.  
Contours of hub neuron frequency error are shown at levels of $.525$, $.53$, ... $.575$ Hz (dark to light gray away from mean).
Dimension of sensitivity $\mathbf{v}_1$ (solid) and robustness $\mathbf{v}_2$ (dashed).
\textbf{F} (Top) The predictive distribution of EPI.
The black and gray dashed lines show the mean and two standard deviations according the emergent property.
(Bottom) Simulations at the starred parameter values.
 }
 \label{fig:STG}
\end{figure}

Before presenting technical details (in the following section), let us understand emergent property inference schematically.
EPI (Fig. \ref{fig:STG}D) takes, as input, the model and the specified emergent property, and as its output, produces the parameter distribution  EPI (Fig. \ref{fig:STG}E).  
This distribution -- represented for clarity as samples from the distribution -- is a parameter distribution that produces the emergent property. 
We can then use this parameter distribution to efficiently generate many parameters producing the emergent property.
Most importantly, the resulting parameter distribution can be efficiently queried to quantify its own structure, which informs the central question of this inverse problem: how do model parameters govern the emergent property?

\subsection{A deep generative modeling approach to emergent property inference} \label{results_dgm}
Emergent property inference (EPI) formalizes the three-step procedure of the previous section with deep probability distributions.
First, as is typical, we consider the model as a coupled set of differential equations.  
In this STG example, the model activity (or state) $\mathbf{x} = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$ is the membrane potential for each neuron, which evolves according to the biophysical conductance-based equation:
\begin{equation} 
\begin{split}
C_m \frac{d\mathbf{x}(t)}{dt} = -h(\mathbf{x}(t); \mathbf{z}) + d\mathbf{B}
 \end{split}
\end{equation} 
where $C_m$=1nF, and $\mathbf{h}$ is a sum of the leak, calcium, potassium, hyperpolarization, electrical, and synaptic currents, all of which have their own complicated dependence on activity $\mathbf{x}$ and parameters $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}]$, and $d\mathbf{B}$ is white gaussian noise \cite{gutierrez2013multiple} (see Section \ref{methods_stg_model} for more detail).

Second, we determine that our model should produce the emergent property of ``intermediate hub frequency" (Figure \ref{fig:STG}C).
\q{stg_ep_text1}{We stipulate that the hub neuron's spiking frequency -- denoted by statistic $\omega_{\text{hub}}(\mathbf{x})$ -- is close to a frequency of 0.55Hz, between that of the slow and fast frequencies.
Mathematically, we define this emergent property with two constraints: that the mean hub frequency is 0.55Hz,}
\begin{equation}
 \mathbb{E}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  0.55\text{Hz}
\end{equation}
\q{stg_ep_text2}{and that the variance of the hub frequency is moderate}
\begin{equation}
\text{Var}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~   0.025^2 \text{Hz}^2.
\end{equation}
In the emergent property of intermediate hub frequency, the statistic of hub neuron frequency is constrained over the distribution of parameters $\mathbf{z}$ and the distribution of the data $\mathbf{x}$ that those parameters produce.
Formally, the emergent property is the collection of these two constraints
\begin{equation}\label{eq:EP_STG}
\mathcal{X}:~~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[\omega_{\text{hub}}(\mathbf{x}; \mathbf{z})\right] = 0.55\text{Hz},~~\text{Var}_{\mathbf{z},\mathbf{x}}\left[\omega_{\text{hub}}(\mathbf{x}; \mathbf{z})\right] =0.025^2\text{Hz}^2.
\end{equation}
In general, an emergent property is a collection of constraints on first-, second- and higher moments of one or more statistics (see Sections \ref{results_RNN}, \ref{results_SC}, and \ref{methods_2DLDS}) that together define the computation.

Third, we perform emergent property inference: we find a distribution over parameter configurations $\mathbf{z}$ that produces the emergent property; in other words, they obey the constraints introduced in Equation \ref{eq:EP_STG}.  
This distribution will be chosen from a family of probability distributions $\mathcal{Q} = \left\{ q_{\bm{\theta}}(\mathbf{z}) : \bm{\theta} \in \Theta \right\}$, defined by a deep neural network \cite{rezende2015variational, dinh2017density, papamakarios2017masked} (Figure \ref{fig:STG}D, EPI box).
Deep probability distributions map a simple random variable $\mathbf{z}_0$ (drawn from an isotropic gaussian) through a deep neural network with weights and biases $\bm{\theta}$ to parameters $\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0)$ of a suitably complicated distribution (see Section \ref{methods_NF} for more details).
\q{results_ME}{Many distributions in $\mathcal{Q}$ will respect the emergent property constraints, so we select the most random (highest entropy) distribution, which is the same choice made in variational bayesian methods (see Section \ref{methods_VI}).}
%\cite{jaynes1957information, elsayed2017structure, loaiza2017maximum, savin2017maximum}
In EPI optimization, stochastic gradient steps in $\bm{\theta}$ are taken such that entropy is maximized, and the emergent property $\mathcal{X}$ is produced (see Section \ref{methods_EPI}).
The inferred EPI distribution is denoted $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$, since it is conditioned upon the implicit datasets of emergent property $\mathcal{X}$ in a manner methodologically equivalent to exemplar data when using variational bayesian approaches.

>>>>

\linenumbers
\section{Abstract}
A cornerstone of theoretical neuroscience is the circuit model: a system of equations that captures a hypothesized neural mechanism.  
Such models are valuable when they give rise to an experimentally observed phenomenon -- whether behavioral or a pattern of neural activity -- and thus can offer insights into neural computation.
The operation of these circuits, like all models, critically depends on the choices of model parameters.
A key step is then to identify the model parameters consistent with observed phenomena: to solve the inverse problem.
To solve challenging inverse problems in neuroscience, statistical inference techniques have been used to infer parameters distributions most likely to have produced neural datasets.
In this work, we present a novel technique, emergent property inference (EPI), that brings the power and versatility of the modern probabilistic modeling toolkit to theoretical neuroscience.
When theorizing circuit models, scientists predominantly focus on reproducing computational properties rather than a particular dataset.
Our method uses deep neural networks to learn parameter distributions with complex structure that produce specific computational properties in circuit models.
This methodology is introduced through a motivational example inferring conductance parameters in a circuit model of the stomatogastric ganglion.
Then, with recurrent neural networks of increasing size, we show that EPI allows precise control over the behavior of inferred parameters, and that EPI scales better in parameter dimension than alternative techniques.
In the remainder of this work, we present novel theoretical findings gained through the examination of complex parametric structure captured by EPI.
In a model of primary visual cortex, we discovered how connectivity with multiple inhibitory subtypes shapes variability in the excitatory population.
Finally, in a model of superior colliculus, we identified and characterized two distinct regimes of connectivity that facilitate switching between opposite tasks amidst interleaved trials, mechanistically characterized each regime using probabilistic tools afforded by EPI, and found conditions where these circuit models reproduce results from optogenetic silencing experiments.
Beyond its scientific contribution, this work illustrates the variety of analyses possible once deep learning is harnessed towards solving theoretical inverse problems.

\section{Introduction}
\q{intro_first}{The fundamental practice of theoretical neuroscience is to use a mathematical model to understand neural computation, whether that computation enables perception, action, or some intermediate processing.  
A neural circuit is systematized with a set of equations -- the model -- and these equations are motivated by biophysics, neurophysiology, and other conceptual considerations \cite{kopell1988coupled,  marder1998biophysics, abbott2008theoretical, wang2010neurophysiological, o2015computational}.
The function of this system is governed by the choice of model \emph{parameters}, which when configured in a particular way, give rise to a measurable signature of a computation.   
The work of analyzing a model then requires solving the inverse problem: given a computation of interest, how can we reason about the space, manifold, or distribution of parameters that give rise to it?  
The inverse problem is crucial for reasoning about likely parameter values, uniquenesses and degeneracies, and predictions made by the model \cite{gutenkunst2007universally, erguler2011practical, mannakee2016sloppiness}.}

@@

\q{intro_data_ep}{Statistical inference, of course, requires quantification of the vague term \emph{computation}.  
In neuroscience, two perspectives are dominant.  
First, often we directly use an \emph{exemplar dataset}: a collection of samples that express the computation of interest, this data being gathered either experimentally in the lab or from a computer simulation.  
While in some sense the best choice given its connection to experiment \cite{paninski2018neural}, some drawbacks exist: these data are well known to have features irrelevant to the computation of interest \cite{niell2010modulation, saleem2013integration, musall2019single}, confounding inferences made on such data.  
Related to this point, use of a conventional dataset encourages conventional data likelihoods or loss functions, which focus on some global metric like squared error or marginal evidence, rather than the computation itself.}

\q{intro_epi}{Alternatively, researchers often quantify an \emph{emergent property} (EP): a statistic of data that directly quantifies the computation of interest, wherein the dataset is implicit.  
While such a choice may seem esoteric, it is not: the above ``gold standard" examples  \cite{hopfield1982neural, sompolinsky1988chaos, olypher2007using, tsodyks1997paradoxical, wong2006recurrent}
all quantify and focus on some derived feature of the data, rather than the data drawn from the model.
An emergent property is of course a dataset by another name, but it suggests different approach to solving the same inverse problem:  here we directly specify the desired emergent property -- a statistic of data drawn from the model -- and the value we wish that property to have, and we set up an optimization program to find the distribution of parameters that produce this computation.  
This statistical framework is not new: it is intimately connected to the literature on approximate bayesian computation \cite{beaumont2002approximate, marjoram2003markov, sisson2007sequential}, parameter sensitivity analyses \cite{raue2009structural, karlsson2012efficient, hines2014determination, raman2017delineating}, maximum entropy modeling \cite{elsayed2017structure, savin2017maximum, mlynarski2020statistical}, and approximate bayesian inference \cite{tran2017hierarchical, gonccalves2019training}; we detail these connections in Section \ref{methods_related}.}

The parameter distributions producing a computation may be thin, curved, bent, or multimodal along various parameter axes and combinations.
It is by capturing and quantifying this complex structure that EPI offers scientific insight.
Traditional approximation families (e.g. mean-field or mixture of gaussians) are limited in the distributional structure they may learn.
To address such restrictions on expressivity, major advancements in machine learning have enabled the use of deep probability distributions as flexible approximating families for such complicated distributions \cite{rezende2015variational, papamakarios2019normalizing} (see Section \ref{methods_NF}).
However, the adaptation of deep probability distributions to the problem of theoretical circuit analysis requires recent developments in deep learning for  constrained optimization \cite{loaiza2017maximum}, and architectural  choices for efficient and expressive deep generative modeling \cite{dinh2017density, kingma2018glow}.
We detail our method, which we call emergent property inference (EPI) in Section \ref{results_dgm}.

Equipped with this method, we demonstrate the capabilities of EPI and present novel theoretical findings from its analysis.
\q{intro_results1}{First, we show EPI's ability to handle biologically realistic circuit models using a five-neuron model of the stomatogastric ganglion \cite{gutierrez2013multiple}: a neural circuit whose parametric degeneracy is closely studied \cite{goldman2001global}.
Then, we show EPI's scalability to high dimensional parameter distributions by inferring connectivities of recurrent neural networks (RNNs) that exhibit stable, yet amplified responses -- a hallmark of neural responses throughout the brain \cite{murphy2009balanced, hennequin2014optimal, bondanelli2019population}.}
\q{intro_results2}{In a model of primary visual cortex \cite{litwin2016inhibitory, palmigiano2020structure}, EPI reveals how the recurrent processing across different neuron-type populations shapes excitatory variability: a finding that we show is analytically intractable.
Finally, we investigated the possible connectivities of superior colliculus that allow execution of different tasks on interleaved trials \cite{duan2018collicular}.
EPI discovered a rich distribution containing two connectivity regimes with different solution classes.
We queried the deep probability distribution learned by EPI to produce a mechanistic understanding of neural responses in each regime. 
Intriguingly, the inferred connectivities of each regime reproduced results from optogenetic inactivation experiments in markedly different ways.}
These theoretical insights afforded by EPI illustrate the value of deep inference for the interrogation of neural circuit models.

@@

Second, once the model is selected, one must specify what the model should produce.
In this STG model, we are concerned with neural spiking frequency, which emerges from the dynamics of the circuit model (Fig. \ref{fig:STG}B).
An emergent property studied by Gutierrez et al. is the hub neuron firing at an intermediate frequency between the intrinsic spiking rates of the fast and slow populations.
This emergent property is shown in Figure \ref{fig:STG}C at an average frequency of 0.55Hz.
Our notion of intermediate hub frequency is not strictly 0.55Hz, but also moderate deviations of this frequency between the fast (.35Hz) and slow (.68Hz) frequencies.
%, which are quantified in the emergent property with variance {0.025$^2$Hz$^2$}.
%To continue our running STG example, one such emergent property is the phenomenon of \emph{network syncing} -- in certain parameter regimes, the frequency of the hub neuron matches that of the fast and slow populations at an intermediate frequency.

%Third, qualitative parameter analysis ensues: since precise mathematical analysis is intractable in this model, a brute force sweep of parameters is done \cite{gutierrez2013multiple}.
%Subsequently, a qualitative description is formulated to describe the different parameter configurations that lead to the emergent property.  
%In this last step lies the opportunity for a precise quantification of the emergent property as a statistical feature of the model. 
% Once we have such a methodology, we can infer a probability distribution over parameter configurations that produce this emergent property.
%Sensitivity analyses \cite{raue2009structural} are also used to identify which parameter contributions have the greatest effect.
%In this last step lies the opportunity for a precise quantification of the emergent property as a statistical feature of the model.

%To infer the STG parameters of intermediate hub frequency with existing methodology, we need an exemplar dataset: experimentally recorded or synthesized.
%In this last step lies the opportunity for a shift away from a dataset-oriented representation of model output towards that of an implicit dataset, where the only structure is the emergent property of interest.
Third, the model parameters producing the emergent property are inferred.
By precisely quantifying the emergent property of interest as a statistical feature of the model, we use EPI to condition directly on this emergent property.
Before presenting technical details (in the following section), let us understand emergent property inference schematically.
EPI (Fig. \ref{fig:STG}D) takes, as input, the model and the specified emergent property, and as its output, returns the parameter distribution (Fig. \ref{fig:STG}E).  
This distribution -- represented for clarity as samples from the distribution -- is a parameter distribution constrained such that the circuit model produces the emergent property. 
Once EPI is run, the returned distribution can be used to efficiently generate additional parameter samples.
Most importantly, the inferred distribution can be efficiently queried to quantify the parametric structure that it captures.
By quantifying the parametric structure governing the emergent property, EPI  informs the central question of this inverse problem: what aspects or combinations of model parameters have the desired emergent property?

\begin{figure}
\begin{center}
\includegraphics[scale=0.8]{figures/fig1/fig1.pdf}
\end{center}
\caption{Emergent property inference (EPI) in the stomatogastric ganglion.  
\textbf{A}. Conductance-based subcircuit model of the STG.
\textbf{B}. Spiking frequency $\omega(\mathbf{x}; \mathbf{z})$ is an emergent property statistic. 
Simulated at $g_{\text{el}} = 4.5$nS and $g_{\text{synA}} = 3$nS.
\textbf{C}. The emergent property of intermediate hub frequency.
Simulated activity traces are colored by log probability of generating parameters in the EPI distribution (Panel E).
\textbf{D}. For a choice of circuit model and emergent property, emergent property inference (EPI) learns a deep probability distribution of parameters $\mathbf{z}$.
\textbf{E}. The EPI distribution producing intermediate hub frequency.
Samples are colored by log probability density.  
Contours of hub neuron frequency error are shown at levels of $.525$, $.53$, ... $.575$ Hz (dark to light gray away from mean).
Dimension of sensitivity $\mathbf{v}_1$ (solid arrow) and robustness $\mathbf{v}_2$ (dashed arrow).
\textbf{F} (Top) The predictions of the EPI distribution.
The black and gray dashed lines show the mean and two standard deviations according the emergent property.
(Bottom) Simulations at the starred parameter values.
 }
 \label{fig:STG}
\end{figure}

\subsection{A deep generative modeling approach to emergent property inference} \label{results_dgm}
Emergent property inference (EPI) formalizes the three-step procedure of the previous section with deep probability distributions \cite{rezende2015variational, papamakarios2019normalizing}.
First, as is typical, we consider the model as a coupled set of noisy differential equations.  
In this STG example, the model activity (or state) $\mathbf{x} = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$ is the membrane potential for each neuron, which evolves according to the biophysical conductance-based equation:
\begin{equation} \label{eq:STG_model1}
\begin{split}
C_m \frac{d\mathbf{x}(t)}{dt} = -h(\mathbf{x}(t); \mathbf{z}) + d\mathbf{B}
 \end{split}
\end{equation} 
where $C_m$=1nF, and $\mathbf{h}$ is a sum of the leak, calcium, potassium, hyperpolarization, electrical, and synaptic currents, all of which have their own complicated dependence on activity $\mathbf{x}$ and parameters $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}]$, and $d\mathbf{B}$ is white gaussian noise \cite{gutierrez2013multiple} (see Section \ref{methods_stg_model} for more detail).

Second, we determine that our model should produce the emergent property of ``intermediate hub frequency" (Figure \ref{fig:STG}C).
\q{stg_ep_text1}{We stipulate that the hub neuron's spiking frequency -- denoted by statistic $\omega_{\text{hub}}(\mathbf{x})$ -- is close to a frequency of 0.55Hz, between that of the slow and fast frequencies.
Mathematically, we define this emergent property with two constraints: that the mean hub frequency is 0.55Hz,}
\begin{equation}\label{eq:EP_STG1}
 \mathbb{E}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  0.55
\end{equation}
\q{stg_ep_text2}{and that the variance of the hub frequency is moderate}
\begin{equation}\label{eq:EP_STG2}
\text{Var}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~   0.025^2
\end{equation}
In the emergent property of intermediate hub frequency, the statistic of hub neuron frequency is constrained over the distribution of parameters $\mathbf{z}$ and the distribution of the data $\mathbf{x}$ that those parameters produce.
This expectation over $\mathbf{x}$ is taken over any randomness in initial state ($\mathbf{x}(t = 0)$) and noise (e.g. $d\mathbf{B}$ in Equation \ref{eq:STG_model1}).
Formally, the emergent property is the collection of these two constraints.
In general, an emergent property is a collection of constraints on statistical moments that together define the computation.

Third, we perform emergent property inference: we find a distribution over parameter configurations $\mathbf{z}$ of models that produce the emergent property; in other words, they satisfy the constraints introduced in Equations \ref{eq:EP_STG1} and \ref{eq:EP_STG2}.  
This distribution will be chosen from a family of probability distributions $\mathcal{Q} = \left\{ q_{\bm{\theta}}(\mathbf{z}) : \bm{\theta} \in \Theta \right\}$, defined by a deep neural network \cite{rezende2015variational, papamakarios2019normalizing} (Figure \ref{fig:STG}D, EPI box).
Deep probability distributions map a simple random variable $\mathbf{z}_0$ (e.g. an isotropic gaussian) through a deep neural network with weights and biases $\bm{\theta}$ to parameters $\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0)$ of a suitably complicated distribution (see Section \ref{methods_NF} for more details).
\q{results_ME}{Many distributions in $\mathcal{Q}$ will respect the emergent property constraints, so we select the most random (highest entropy) distribution, which is the same choice commonly made in variational bayesian methods (see Section \ref{methods_VI}).}
%\cite{jaynes1957information, elsayed2017structure, loaiza2017maximum, savin2017maximum}
In EPI optimization, stochastic gradient steps in $\bm{\theta}$ are taken such that entropy is maximized, and the emergent property $\mathcal{X}$ is produced (see Section \ref{methods_EPI}).
We then denote the inferred EPI distribution as $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$, since the structure of the learned parameter distribution is determined by weights and biases $\bm{\theta}$, and this distribution is conditioned upon emergent property $\mathcal{X}$.
